\documentclass[a4paper]{article}
%\documentclass[13pt,a4paper,twoside,titlepage]{article}
\usepackage[ngerman, english]{babel}
\usepackage[utf8]{inputenc}
\usepackage{latexsym,amssymb,amsmath,amstext}
\usepackage{graphicx}
\usepackage{float}
\usepackage[backend=bibtex, sorting=none]{biblatex}
\usepackage{amsmath}
\usepackage{ dsfont }
\usepackage{csquotes}
\usepackage{slashed}
\usepackage{subfigure}
\usepackage{listings}
\usepackage{subfigure}
\usepackage{verbatim}
\usepackage{xcolor}
\usepackage{multicol}
%\usepackage{feynmf}
\usepackage{abstract}
\usepackage{bbold}
%\usepackage[a4paper, left=2cm, right=2cm, top=2cm, bottom=3cm, bindingoffset=10mm]{geometry}
\usepackage[compat=1.1.0]{tikz-feynman}
%\usepackage[onehalfspacing]{setspace}
\usepackage{tikz}
\usepackage{pgfplots}
\lstset{
  breaklines=true,
  postbreak=\mbox{\textcolor{red}{$\hookrightarrow$}\space},
}

\newcommand{\Tr}{\operatorname{Tr}}
\newcommand{\note}[1]{{\color{red}[\textbf{#1}]}}
\newcommand{\diff}{\ensuremath{\mathrm{d}}}

\bibliography{refs}

\begin{document}

\section{Introduction}
In this brief report we will report on the patras workshop 2021.
First we describe the scope of the workshop, then three talks on two different topics.
Finally we describe the online-format and conclude by reviewing the workshop from our own perspective.

\section{The Patras-Workshop}
The 16th Patras workshop on axions, WIMPs and WISPs has been held online from the 14.6.2021 to the 18.6.2021 and was organized by the university of Trieste.
Its main focus was on axions especially axion detection experiments.
Other topics included dark photons, dark matter modeling and
detection,  axion stars,  bounds on different dark matter candidates (mostly axions).

\section{Talks}
First let us talk about topological cosmic axion strings.
Axion strings are one dimensional topological defects in the UV completing
(PQ) field of the axion. The field winds the field space on a loop
in real space. Contracting the loop to a single point, while this ``winding number'' stays the same, yields a point, where the axion field is undefined. In three dimensions this can happen along a one-dimensional curve - the axion string.
During the cosmological evolution, the energy of those strings is transferred into axions.
Hence the axion dark matter relic density is influenced by this process.
The qualitative amount (significant or sub-leading) contribution as well as the precise quantitative amount
of axions produced by cosmic strings was for the last decades and still is, a topic of debate  \cite{Gorghetto_2018}. \\


\noindent
There were two talks given on axion strings.
The first was a short talk by Asier Lopez-Eiguren on ``Approach to scaling in axion string networks'' based on their paper with the same title \cite{Hindmarsh_2021}.
They performed numerical simulations of the string network and compared it to an effective
model. As the result they found that the dynamics is well described by the equations of the effective theory. That means that the `logarithmic violating'' observed by other groups is only a transient effect.
Their result is not fully conclusive as they couldn't simulate the string network to the point, where
the transient effect vanishes. \\

\noindent
The second talk was given by Marco Gorghetto on gravitational wave signals
from axions strings. They found that the signal could be described by an effective theory
and found new bounds on the axion parameter space from gravitational wave detectors \cite{Gorghetto_2021}. \\

\noindent
Regarding the topic of my masters thesis, only one talk was about baryogenesis.
This talk was about axion-quark-nuggets by Ariel Zhitnitsky. This model aims to explain
dark matter and baryogenesis at the same time. The relic densities of the dark matter
and baryonic matter are related in this model and hence it can explain why the dark matter
and the baryon densities are of the same order of magnitude. In this theory
a new state of QCD matter is proposed. The nuggets also come in form of anti-quark-nuggets.
These nuggets contain the ''missing``anti-matter and also serve as the dark matter particles.
They are stabilized by an axion domain wall. How this mechanism works i.e. how the nuggets
form and why the anti-matter is inside the nuggets and not the matter is beyond my understanding from
the talk.
The focus of the talk was one experimental detection. It was based on the work by the speaker as a
co-author of ref. \cite{Budker_2020}. A review paper by the speaker is ref. \cite{Zhitnitsky_2021}.


\section{The Online-Conference Format}
Let us now discuss the format of the conference.
Due to the ongoing COVID-19 pandemic the conference was held online using Zoom.
There were 25 minute talks as well as 5 minute presentations that replaced poster presentations.
The presentations were given with almost no breaks in between and little room for questions (only one or two
questions here allowed). This lead to the conference always going over time.

\section{Review}
I will comment on the online format from my perspective.
The positive side of the online format is the easier and cheaper (in this case completely free)
attendance, because of lesser or non attendance fees and no travel and accommodation costs.
This makes it easier, in particular for students, to attend. As well as lower environmental impact from traveling.
The flip side of this is that the barrier for real networking is much higher.
All questions are asked in front of the hole audience. This includes the
poster presentations. The breakout rooms for the breaks were only used by a few people, who apparently already knew
each other.
All this is a bit intimidating.
The topics of the talks are quite advanced for a student.
This is of course always the case, online or not. But in the online case no opportunity is given to have
conversations in smaller groups, which I imagine to be more beginner friendly.
In the end it was interesting to hear about various current research topics but
I didn't really do any networking.
I also feel like watching online talks and starring on screens is more tiring than going to in-person talks.


\newpage
\printbibliography
\end{document}

